---
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(ggplot2)
```


```{r}
cytof = read_csv("https://jfukuyama.github.io/teaching/stat670/notes/cytof_one_experiment.csv")
```


```{r}
head(cytof)
```


# Q.1)

# Answer

1. Converting the CyTOF dataset from wide form to a longer form
2. We first check the column names and number of col names

```{r}
colnames(cytof)
```
```{r}
ncol(cytof)
```




```{r}
cytof_long= cytof %>%
  pivot_longer(c(colnames(cytof)),names_to = "markers",values_to = "values")
```


3. This is the data we get after converting to longer form

```{r}
head(cytof_long)
```

4. Now we plot the faceted plots for distributions of all markers using density plot and
also we created a histogram plot for the same

```{r}
marker_gg = ggplot(cytof_long,aes(x=values)) + geom_density()
marker_gg + facet_wrap(~ markers , nrow = 5) + ggtitle("Distribution of markers")
```


```{r}
marker_gg = ggplot(cytof_long,aes(x=values)) + geom_histogram()
marker_gg + facet_wrap(~ markers , nrow = 5) + ggtitle("Distribution of markers")
```

5. From the above faceted plots, we see the folloing:
  a. Markers like CD56 and CD69 show similar distribution.
  b. When we check the mean and median values for all the distribution using summary method(refer Q.3),we see that there are very few markers who follow a normal distribution.
  c. Mostly the markers are right-skewed and left-skewed based on the values of mean and median.
  d. The standard deviation for most of the markers is around 1 or in most cases greater than 1. So the distribution is quite spread out out.
  e. Also we see that markers KRDL1 and CD8 are also similar in distribution.




# Q.2)

# Answer:

1. Here from looking at the graph, we took two sets of markers
2. For the first set, we took the markers CD69 and CD56 because they almost are 
similar in distribution.


```{r}
CD69 <- filter(cytof_long, markers == "CD69") %>%  pull(values)
```


```{r}
CD56 <- filter(cytof_long, markers == "CD56") %>%  pull(values)
```


```{r}
qqplot(x=CD69,y=CD56)
abline(h=median(CD69),col="blue")
abline(v=median(CD56),col="red")
title("Q-Q plot for Markers: CD69 and CD56")
```
3. Here since the distributions are almost similar, the points for the quartiles lie
almost in a straight line except for the region near the origin.
4. Near this region, the quartiles seem to increase exponentially
5. We then plot the median for both the marker and we see that the median value for marker
CD56 is around 4.5 and the median for marker CD69 is around 5.7
6. Since median values of both the markers are different and shape of distribution is almost
the same, this is a simple shift.
7. For the parts around -0.6 to 0 the quartiles are increasing and then after 0 the quartiles fall in a line.


```{r}
CD57 <- filter(cytof_long, markers == "CD57") %>%  pull(values)
```

```{r}
qqplot(x=CD56,y=CD57)
abline(h=median(CD57),col='blue')
abline(v=median(CD56),col='red')
title("Q-Q plot for Markers: CD57 and CD56")

```
8. For the second set, we took the markers CD57 and CD56 because they are 
dissimilar in distribution.
9. Here since the distributions are not similar, the points for the quartiles do not lie
in a straight line.
10. We check the median values for the markers CD57 and CD56.
11. The median value for CD57 is around 3 and median value for CD56 is around 5.7
12. Since the values of medians for both the markers is different and the shape of distribution is also different the shift for this distribution seems to be complex
13. For the parts around -0.6 to 0, there seems to be outliers in that region, after around 0.2 till 4, the quartiles are steady and after 4.5 onwards the quartiles are exponentially increaing.


# Q.3)

# Answer:
1. We first computed the statistics on the basis of all markers
2. From this, we get the value of mean,median,1st quartile,3rd quartile,min and max

```{r}
by(cytof_long,cytof_long$markers,summary)
```
3. We also computed other statistical data which was not previously computed using the summary method 
4. Here we computed the measures of spread of distributions for all the markers by first
finding the standard deviation and then the IQR
5. We see that for most of the markers, the standard deviation is greater than 1 and very close to 1 so we can say that the data is more spread out for the markers.

```{r}
cytof_long %>% group_by(markers) %>% summarise(sd(values)) 
```

```{r}
cytof_long %>% group_by(markers) %>% summarise(IQR(values)) 
```

6. Now for plotting the statistical data, we have chosen box plot.
7. From box plot we get to see the data for the median, the 1st and 3rd quartile, the min and the max values for all the markers
8. As compared to the full distribution we plotted before, this gives a fairly better statistical measure for the data as opposed to the histogram where we have to estimate the values of mean and spread.


```{r}
ggplot(cytof_long) +
  aes(x = markers, y =values) +
  geom_boxplot() + coord_flip()
```

